The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-04-17 paper/referee-reply.tex

\documentclass[12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[spanish]{babel}
\usepackage{amsmath}
\usepackage{graphicx}

\title{Response to the referee}

\begin{document}

\date{}
\maketitle

\section*{Outflow + Rotation}

We want to study the pure effect of rotation on the spectra, off course it would
be interesting to study the effect of rotation on systems with outflows, this paper
is under preparation by M.C Remolina-Gutierrez et al.

\section*{Details}

\subsection*{Introduction}

This changes were performed.

\subsection*{Fig.1}

\begin{itemize}
\item Description about how we build Fig.1 (Done)
\item Fig 1. remove x notation (Done)
\end{itemize}

\subsection*{Fig.2}

Explanation (Done)\\
We don't fit a gaussian , we measure the width by finding
the closest points to the half maximum in our histogram numerically.

\subsection*{Fig.3}

\subsection*{Fig4 and 6}

Waiting for the cluster in germany!

\subsection*{Fig5}

\end{document}

### 2014-04-21 paper/ms.tex
Fig2. comments about the FWHM
2d Histograms of spectra (Homgeneous, Central)
sec 3 a little bit of text

A quantitative description of the \ly line has been reached through
Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. In the last
two decades these studies have become popular due to the
availability of computing power. Early into the 21st century, the first
studies focused on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}; Later on, the effects of clumpy media
\citep{Hansen06} and expanding/contracting shell/spherical geometries started to
be studied \citep{Verhamme06,Dijkstra06,Ahn2014}. Similar codes have applied
these results to semi-analytic models of galaxy formation \citep{Orsi12, Garel2012} and
results of large hydrodynamic simulations
\citep{CLARA,Forero12,Behrens13}. Recently, Monte Carlo codes have also
been applied to the results of high resolution hydrodynamic
simulations of individual
galaxies\citep{Laursen09,Barnes11,Verhamme12,Yajima12}. Meanwhile, recent
developments have been focused on the systematic study of clumpy
outflows \citep{DijkstraKramer}and anisotropic velocity configurations
\citep{Zheng2013}.

@@

For each photon in the simulation we have its initial position inside
the sphere, direction of propagation $\hat{k}_{in}$ and reduced
frequency $x_{in}$. The photon's propagation stops once they cross the
surface of the sphere. At this point we store the position, the outgoing direction
of propagation $\hat{k}_{out}$ and the reduced frequency $x_{out}$. We
define the angle $\Theta$ by $\cos\Theta = \hat{k}_{\rm obs}\cdot
\hat{k}_{\rm out}$, that is the angle of the outgoing photon with
respect to the directions towards the observer. Following
\cite{Zheng2013} we make the study of the anisotropic emission in
terms of this angle.

@@

@@

The central result of this paper is summarized in Figure
\ref{fig:differentvelocities}. It shows the clear
impact of rotation on the morphology of the emergent \ly line,
This profiles were obtained integrating the flux of photons 
 over all directions, this means we made histograms of the 
 numbers of photons escaping the galaxy with a given frequency
The panel focus on the results for all of our symmetric models
 $\tau_{H}=(10^{5}, 10^{6}, 10^{7}$), showing that the
influence of rotation is present both when the photons are either
homogeneously or centrally initialized over the gas volume.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{f2.eps}
\end{center}
  \caption{FWHM for the non-dusty models as a function of
    rotational velocity $V_{\rm max}$. The left panel shows the
    results in velocity units while the right panel normalizes
    the data by the FWHM in the static case. 
    Continuous (dashed) lines  correspond to central (homogeneous)
    source distributions. The straight lines represent the fit to
    the data using the expression in Eq. (\ref{eq:fwhm}).
    \label{fig:widthvsvelocity}} 
\end{figure*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f3.eps}
\end{center}
\caption{Position of the peak maxima as a function of rotational
  velocity $V_{\rm max}$. Continuous (dashed) lines correspond to
  central (homogeneous) source distributions. A value of $x_{\rm
    max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}  
\end{figure}

@@

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parametrize the dependency of the line width with
$V_{\rm max}$ as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.89 \pm 0.04$ and $\lambda_{\rm h}=
1.98\pm 0.06$ respectively.

@@

\subsection{Anisotropy and off-centered emission}
\label{sec:off-center}

@@

Are these features observable with current instrumental capabilities?
As an order of magnitude estimate we can consider that the width of
each one of the triple peaks is close to $1/3$ of the FWHM of the line
at a given $V_{\rm max}$. For the cases we have studied in this
subsection those values would be close to $50\kms$, $100\kms$ and
$200\kms$ for optical depths of $10^5$, $10^6$, and $10^7$
respectively. At a redshift of $z\sim 2-3$ each one of these widths
is multiplied by factors of $z+1\sim 3-4$. Considering a nominal value
for the resolution of $\Delta v=200\kms$ for a Near-IR spectrograph capable of observing
those galaxies \citep{Kulas12}, we find that only the case of triple
peaked emission for the most massive object $\tau_{\rm H}=10^{7}$,
$V_{\rm max}=200\kms$ would have a chance to be spectrally
resolved. We note that observations of triple peaked systems with
higher resolution spectrographs have been reported  by
\cite{Chonis2013}, however the large velocity off-sets ($\sim
1000$\kms) from the systemic redshift favor the hypothesis of an
interacting system, including emission from shocked and fluorecing
gas, as the origin of the triple peaked emission.

>>>>

A quantitative description of the \ly line has been reached through
Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. In the last
two decades these studies have become popular due to the
availability of computing power. Early into the 21st century, the first
studies focused on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}; Later on, the effects of clumpy media
\citep{Hansen06} and expanding/contracting shell/spherical geometries started to
be studied \citep{Verhamme06,Dijkstra06,Ahn2014}. Similar codes have applied
these results to semi-analytic models of galaxy formation \citep{Orsi12, Garel2012} and
results of large hydrodynamic simulations
\citep{CLARA,Forero12,Behrens13}. Recently, Monte Carlo codes have also
been applied to the results of high resolution hydrodynamic
simulations of individual
galaxies \citep{Laursen09,Barnes11,Verhamme12,Yajima12}. Meanwhile, recent
developments have been focused on the systematic study of clumpy
outflows \citep{DijkstraKramer}and anisotropic velocity configurations
\citep{Zheng2013}.

@@

For each photon in the simulation we have its initial position inside
the sphere, direction of propagation $\hat{k}_{in}$ and reduced
frequency $x_{in}$. The photon's propagation stops once they cross the
surface of the sphere. At this point we store the position, the outgoing direction
of propagation $\hat{k}_{out}$ and the reduced frequency $x_{out}$. We
define the angle $\theta$ by $\cos\theta = \hat{k}_{\rm obs}\cdot
\hat{k}_{\rm \hat{k}}$, that is the angle of the line of sight with
respect to the direction of the angular velocity. Following the notation
implemented in \cite{Zheng2013}. %we make the study of the anisotropic emission in
%terms of this angle.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{spectraCentral.eps}
\end{center}
\caption{Central.
    \label{fig:CentralSpec}}  
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{spectraHOM.eps}
\end{center}
\caption{Homogeneous
    \label{fig:HOMSpec}}  
\end{figure*}

@@

The central results of this paper is summarized in Figures
\ref{fig:CentralSpec} and \ref{fig:HOMSpec}. These figures shows
that rotation fix a preferred axis for both photons distributions, 
this implies that our systems doesn't have spherical symmetry and
we have to take into account the viewing angle $\theta$ for all
of our posterior analysis.

Figure \ref{fig:differentvelocities} shows the clear
impact of rotation on the morphology of the emergent \ly line,
This profiles were obtained integrating the flux of photons 
 over the angle $\theta = \pi/2$, this means we made histograms of the 
 numbers of photons escaping the galaxy with a given frequency
The panel focus on the results for all of our symmetric models
 $\tau_{H}=(10^{5}, 10^{6}, 10^{7}$), showing that the
influence of rotation is present both when the photons are either
homogeneously or centrally initialized over the gas volume.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{widths.eps}
\end{center}
  \caption{FWHM for the non-dusty models as a function of
    rotational velocity $V_{\rm max}$. The left panel shows the
    results in velocity units while the right panel normalizes
    the data by the FWHM in the static case. 
    Continuous (dashed) lines  correspond to central (homogeneous)
    source distributions. The straight lines represent the fit to
    the data using the expression in Eq. (\ref{eq:fwhm}).
    \label{fig:widthvsvelocity}} 
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{maxima.eps}
\end{center}
\caption{Position of the peak maxima as a function of rotational
  velocity $V_{\rm max}$. Continuous (dashed) lines correspond to
  central (homogeneous) source distributions. A value of $x_{\rm
    max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}  
\end{figure*}

@@

We use the full width at half maximum (FWHM) to quantify the line
broadening. For this we measure the width by finding
the closest points to the half maximum in our histogram numerically. 
 Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parametrize the dependency of the line width with
$V_{\rm max}$ as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.89 \pm 0.04$ and $\lambda_{\rm h}=
1.98\pm 0.06$ respectively.

@@

\iffalse
\subsection{Anisotropy and off-centered emission}
\label{sec:off-center}

@@

Are these features observable with current instrumental capabilities?
As an order of magnitude estimate we can consider that the width of
each one of the triple peaks is close to $1/3$ of the FWHM of the line
at a given $V_{\rm max}$. For the cases we have studied in this
subsection those values would be close to $50\kms$, $100\kms$ and
$200\kms$ for optical depths of $10^5$, $10^6$, and $10^7$
respectively. At a redshift of $z\sim 2-3$ each one of these widths
is multiplied by factors of $z+1\sim 3-4$. Considering a nominal value
for the resolution of $\Delta v=200\kms$ for a Near-IR spectrograph capable of observing
those galaxies \citep{Kulas12}, we find that only the case of triple
peaked emission for the most massive object $\tau_{\rm H}=10^{7}$,
$V_{\rm max}=200\kms$ would have a chance to be spectrally
resolved. We note that observations of triple peaked systems with
higher resolution spectrographs have been reported  by
\cite{Chonis2013}, however the large velocity off-sets ($\sim
1000$\kms) from the systemic redshift favor the hypothesis of an
interacting system, including emission from shocked and fluorecing
gas, as the origin of the triple peaked emission.
\fi
